---
title: "Covid-Liver"
author: "Krish Shah"
date: "`r Sys.Date()`"
output: html_document
---
#Load libraries
```{r}
library(survival)
library(naniar)
library(visdat)
library(ggplot2)
library(mice)
library(missForest)
library(readr)
library(VIM)
library(dplyr)
library(corrplot)
library(survminer)
library(pROC)
library(caret)
library(glmnet)
```


```{r}
df <- read.csv('covid-liver.csv')
head(df)
```


```{r}
dim(df)
summary(df)
```


```{r}
character_columns <- sapply(df, is.character)

# Extract names of character columns
character_column_names <- names(character_columns)[character_columns]
print(character_column_names)
```


```{r}
# Convert character columns to factors
df[character_column_names] <- lapply(df[character_column_names], as.factor)
head(df)
```

#Little's MCAR Test
```{r}
mcar_test(df)
```

#Missing Data Visualisation



```{r}
vis_miss(df) +
  theme(axis.text.x =  element_text(angle = 90))
```


```{r}
library(VIM)
aggr_plot <- aggr(df, col=c('navyblue','red'), numbers=TRUE, sortVars=TRUE, 
                  labels=names(df), cex.axis=.7, gap=3, ylab=c("Histogram of missing data","Pattern"))
```


```{r}
gg_miss_upset(df)
# gg_miss_span(High_Level_Indices,Electoral,span_every = 1000)
```
```{r}
n_var_miss(df)
```
```{r}
gg_miss_upset(df, nsets = n_var_miss(df))
```
```{r}
gg_miss_upset(df,
              nsets = 10,
              nintersects = NA)


```

```{r}
gg_miss_case(df)		     #Plot the number of missings per case (row)
gg_miss_case_cumsum(df)	     #Plot of cumulative sum of missing for cases
gg_miss_fct(df, Cancer)		     #Plot the number of missings for each variable, broken down by a factor
gg_miss_var(df,show_pct = TRUE)   #Plot the number of missings for each variable
gg_miss_var_cumsum(df)	 


```
```{r}
ggplot(df,
        aes(x = Size,
        y = Age)) +
        geom_miss_point() +
        facet_wrap(~Treatment_grps)

ggplot(df,
        aes(x = Size,
        y = Age)) +
        geom_miss_point() +
        facet_wrap(~Cancer)

ggplot(df,
        aes(x = Size,
        y = Age)) +
        geom_miss_point() +
        facet_wrap(~Year)

ggplot(df,
        aes(x = Size,
        y = Age)) +
        geom_miss_point() +
        facet_wrap(~HCC_BCLC_Stage)

ggplot(df,
        aes(x = Size,
        y = Survival_fromMDM)) +
        geom_miss_point() +
        facet_wrap(~Alive_Dead)
```


```{r}
ggplot(df, aes(fill=Alive_Dead, y=Survival_fromMDM, x=Treatment_grps)) +
  geom_bar(position='dodge', stat='identity') +
  ggtitle('Survival from Multidisciplinary Meeting by Treatment Groups') +
  xlab('Treatment_groups') +
  ylab('Survival_fromMDM (in Months)') +
  scale_fill_manual('Event', values=c('coral2','blue'))
```

#Drop Columns
```{r}
#We drop columns with more than 32% missing values 
sapply(df, function(x) sum(is.na(x)))
columns_to_drop <- c("Bleed","ICC_TNM_Stage" , 
                     "Type_of_incidental_finding","Surveillance_effectiveness",
                     "Mode_of_surveillance_detection","Time_diagnosis_1st_Tx" ,
                     "Date_incident_surveillance_scan","Time_MDM_1st_treatment",
                     "Time_decisiontotreat_1st_treatment","Months_from_last_surveillance")

column_indices_to_drop <- which(names(df) %in% columns_to_drop)
df <- df[, -column_indices_to_drop]
head(df)
```
```{r}
dim(df)
```

# Imputation:


```{r}
Testing <-  df[complete.cases(df), ]
Training <- prodNA(Testing, noNA = 0.3)
dim(Training)
```


```{r}
# Run mice with different methods
imp1 <- mice(Training, method = "pmm", m = 5, maxit = 10, seed = 500) 
summary(imp1)
```


```{r}
xyplot(imp2, Size ~ Survival_fromMDM + Age, pch=18, cex=1)
```
```{r}
imp2 <- mice(Training, method = "rf", m = 5, maxit = 10, seed = 500)
summary(imp2)
```


```{r}
missforest_data_df <- missForest(Training, verbose = TRUE)
missforest_data_df <- as.data.frame(missforest_data_df$ximp)

```


```{r}
rf_data <- complete(imp2,3)
pmm_data <- complete(imp1, 3)
rf_data <- as.data.frame(rf_data)
pmm_data <- as.data.frame(pmm_data)

```


```{r}
# Create a data frame to store results
results <- data.frame(Method = character(0), RMSE = numeric(0), MAE = numeric(0))

# Define a function to calculate RMSE and MAE
calculate_metrics <- function(predictions, actual) {
  rmse <- sqrt(mean((predictions - actual)^2))
  mae <- mean(abs(predictions - actual))
  return(c(RMSE = round(rmse, 2), MAE = round(mae, 2)))
}


# MissForest approach
method <- "MissForest"
# Define X and y for MissForest approach
X_missforest <- as.matrix(model.matrix(~ . - 1, data = missforest_data_df))
y <- as.vector(missforest_data_df$Size)
# Run cross-validated glmnet with optimal lambda
cv_fit_missforest <- cv.glmnet(X_missforest, y)
# Get the lambda value that minimizes the mean cross-validated error
optimal_lambda_missforest <- cv_fit_missforest$lambda.min
# Fit glmnet with the optimal lambda value
fit_optimal_lambda_missforest <- glmnet(X_missforest, y, alpha = 0.5, lambda = optimal_lambda_missforest)
# Predict on the training set
predictions_missforest <- predict(fit_optimal_lambda_missforest, newx = X_missforest)
# Calculate RMSE and MAE
metrics_missforest <- calculate_metrics(predictions_missforest, y)
results <- rbind(results, c(method, metrics_missforest))

# PMM approach 
method <- "Predictive Mean Matching"

X_pmm <- as.matrix(model.matrix(~ . - 1, data = pmm_data))
y <- as.vector(pmm_data$Size)
# Run cross-validated glmnet with optimal lambda
cv_fit_pmm <- cv.glmnet(X_pmm, y)
# Get the lambda value that minimizes the mean cross-validated error
optimal_lambda_pmm <- cv_fit_pmm$lambda.min
# Fit glmnet with the optimal lambda value
fit_optimal_lambda_pmm <- glmnet(X_pmm, y, alpha = 0.5, lambda = optimal_lambda_pmm)
# Predict on the training set
predictions_pmm <- predict(fit_optimal_lambda_pmm, newx = X_pmm)
# Calculate RMSE and MAE
metrics_pmm <- calculate_metrics(predictions_pmm, y)
results <- rbind(results, c(method, metrics_pmm))

# Random Forest imputation approach 

method <- "Random Forest Imputation"
# Define X and y for Random Forest imputation approach
X_rf <- as.matrix(model.matrix(~ . - 1, data = rf_data))
y <- as.vector(rf_data$Size)
# Run cross-validated glmnet with optimal lambda
cv_fit_rf <- cv.glmnet(X_rf, y)
# Get the lambda value that minimizes the mean cross-validated error
optimal_lambda_rf <- cv_fit_rf$lambda.min
# Fit glmnet with the optimal lambda value
fit_optimal_lambda_rf <- glmnet(X_rf, y, alpha = 0.5, lambda = optimal_lambda_rf)
# Predict on the training set
predictions_rf <- predict(fit_optimal_lambda_rf, newx = X_rf)
# Calculate RMSE and MAE
metrics_rf <- calculate_metrics(predictions_rf, y)
results <- rbind(results, c(method, metrics_rf))


colnames(results) <- c("Method", "RMSE", "MAE")

# Print or display the results
print(results)



```

```{r}
densityplot(imp1)
densityplot(imp2)



```

```{r}

#Impute the entire dataset
impute_data <- missForest(df, verbose = TRUE)
df <- as.data.frame(impute_data$ximp)
```

#  1). check whether cancer effected people are more in Pandemic compare to Prepandemic

Ho(null-hypothesis): cancer effected during pandemic = cancer effected during prepandemic

Ha(alternative-hypothesis): cancer effected during pandemic != cancer effected during prepandemic

```{r}
contingency_table <- table(df$Year, df$Cancer)
print(contingency_table)
```
```{r}
proportions <- prop.table(contingency_table, margin = 1)
print(contingency_table)
# Step 3: Conduct a two-sample proportion test
test_result <- prop.test(c(120, 190), c(120 + 64, 190 + 76), correct = FALSE)
print(test_result)

# Step 4: Visualize the contingency table using a bar plot
barplot(as.matrix(contingency_table), beside = TRUE, legend.text = FALSE, col = c("blue", "orange"), main = "Cancer Cases by Year", xlab = "Year", ylab = "Frequency")
legend("topleft",    # Position of the legend
       legend = c("Pandemic", "Prepandemic"),   # Legend labels
       fill = c("blue", "orange"))  

```

#does cancer effects gender?

Ho(null-hypothesis): male having cancer = female having cancer

Ha(alternative-hypothesis) male having cancer != female having cancer
```{r}

# Step 1: Create the contingency table
contingency_table <- table(df$Gender, df$Cancer)
print(contingency_table)

# Step 2: Calculate proportions
proportions <- prop.table(contingency_table, margin = 1)

# Step 3: Conduct a two-sample proportion test
# Perform proportion test
test_result <- prop.test(c(62, 244), c(52 + 62, 87 + 244), correct = FALSE)
print(test_result)

# Step 4: Visualize the contingency table using a bar plot
barplot(as.matrix(contingency_table), beside = TRUE, legend.text = FALSE, col = c("blue", "orange"), main = "Cancer Cases by Gender", xlab = "Cancer", ylab = "Frequency")

# Add legend
legend("topleft",    # Position of the legend
       legend = c("Female", "Male"),   # Legend labels
       fill = c("blue", "orange"))

# Add title for legend
title(legend = "Cancer")

 
```
```{r}
# Step 1: Create the contingency table
contingency_table <- table(df$Gender, df$Year)
print(contingency_table)
```


```{r}
# Filter the data to include only observations during the pandemic period
pandemic_data <- subset(df, Year == "Pandemic")
dim(pandemic_data)
# Create a contingency table of gender and cancer flag for the pandemic period
contingency_table <- table(pandemic_data$Gender, pandemic_data$Cancer)
print(contingency_table)

# Perform a two-sample proportion test
test_result <- prop.test(c(contingency_table[1, 2], contingency_table[2, 2]), 
                          c(rowSums(contingency_table)), correct = FALSE)
print(test_result)


```
```{r}
barplot(as.matrix(contingency_table), beside = TRUE, legend.text = FALSE, col = c("blue", "orange"), main = "Cancer Cases by Gender during Pandemic", xlab = "Cancer", ylab = "Frequency")

# Add legend
legend("topleft",    # Position of the legend
       legend = c("Female", "Male"),   # Legend labels
       fill = c("blue", "orange"))



```


```{r}
t_test <- t.test(Size ~ Year, data = df)
print(t_test)
```
```{r}
# Create a contingency table of BCLC stages by year
contingency_table <- table(df$HCC_BCLC_Stage, df$Year)

# Display the contingency table
print(contingency_table)

# Perform a chi-square test of independence
# Perform Fisher's exact test
fisher_test <- fisher.test(contingency_table)
print(fisher_test)


```

```{r}
ggplot(df, aes(x = Year, y = Size, fill = Year)) +
  geom_boxplot() +
  labs(title = "Comparison of Tumor Size by Pandemic Period",
       x = "Year",
       y = "Tumor Size") +
  scale_fill_manual(values = c("Prepandemic" = "blue", "Pandemic" = "orange")) +  # Set colors for pandemic periods
  theme_minimal()
```

```{r}
table_presentation <- table(df$Mode_Presentation, df$Year)
# Perform chi-squared test
chisq_test <- chisq.test(table_presentation)
print(chisq_test)

ggplot(df, aes(x = Mode_Presentation, fill = Year)) +
  geom_bar(position = "dodge", color = "black") +
  labs(title = "Mode of Presentation by Year",
       x = "Mode of Presentation",
       y = "Count") +
  scale_fill_manual(values = c("Prepandemic" = "blue", "Pandemic" = "orange")) +  # Set colors for each year
  theme_minimal()
```

```{r}
#Convert to Numerical values
df1<- df
df1$Cancer<- as.numeric(factor(df1$Cancer))
df1$Cirrhosis <- as.numeric(factor(df1$Cirrhosis))
df1$Surveillance_programme<- as.numeric(factor(df1$Surveillance_programme))
df1$HCC_TNM_Stage<- as.numeric(factor(df1$HCC_TNM_Stage))
df1$Etiology<- as.numeric(factor(df1$Etiology))
df1$Year<-as.numeric(factor(df1$Year))
df1$Mode_Presentation<-as.numeric(factor(df1$Mode_Presentation))
df1$Gender<-as.numeric(factor(df1$Gender))
df1$HCC_BCLC_Stage<-as.numeric(factor(df1$HCC_BCLC_Stage))
df1$Treatment_grps<-as.numeric(factor(df1$Treatment_grps))
df1$Alive_Dead<-as.numeric(factor(df1$Alive_Dead))
df1$Prev_known_cirrhosis<-as.numeric(factor(df1$Prev_known_cirrhosis))
```


```{r}
pie_data <- df %>%
  group_by(Year, Alive_Dead) %>%
  summarise(count = n()) %>%
  mutate(percent = count / sum(count) * 100)
# Define custom colors
custom_colors <- c("Dead" = "coral", "Alive" = "blue")

# Plot pie charts
pie_charts <- list()
for (year in unique(df$Year)) {
  plot_data <- pie_data %>%
    filter(Year == year)
  
  pie_charts[[year]] <- ggplot(plot_data, aes(x = "", y = percent, fill = Alive_Dead)) +
    geom_bar(stat = "identity") +
    coord_polar("y", start = 0) +
    labs(title = paste("", year)) +
    scale_fill_discrete(labels = c("Dead", "Alive")) + scale_fill_manual(values = custom_colors) +  # Set custom colors # Customize the legend labels
    theme_void() +
    theme(legend.position = "bottom")
}

# Arrange the pie charts
gridExtra::grid.arrange(grobs = pie_charts, ncol = 2)

```
```{r}
ggplot(data = df1, aes(x = Age, y = Size, color = Cancer)) +
  geom_point() +
  labs(title = "Relationship between Age and Size by Cancer Status",
       x = "Age", y = "Size") +
  theme_minimal()

boxplot(df1, col = "blue")


#Correlation Analysis

cor_matrix<-cor(df1)
corrplot(cor_matrix,tl.pos ='lt',tl.col="black", tl.cex=0.55,method='circle')
```

```{r}
ggplot(df, aes(x = Year, fill = Year)) +
  geom_bar(color = "black", width = 0.6) +  # Adjust the width as needed (default is 0.9)
  labs(title = "Count of Liver Cancer Cases by Year",
       x = "Year",
       y = "Count") + 
  scale_fill_manual(values = c("Prepandemic" = "blue", "Pandemic" = "orange")) +  # Set colors for each year
  theme_minimal() +
  theme(legend.position = "none")

ggplot(df, aes(x = Gender, fill = Year)) +
  geom_bar(position = "dodge", color = "black") +
  labs(title = "Gender Distribution by Pandemic Period",
       x = "Gender",
       y = "Count") +
  scale_fill_manual(values = c("Prepandemic" = "blue", "Pandemic" = "orange")) +  # Set colors for pandemic periods
  theme_minimal()

```
## Survival Analysis
```{r}
df$Alive_Dead<-as.numeric(factor(df$Alive_Dead))

# Create the survival fit object
fit.surv <- survfit(Surv(df$Survival_fromMDM, df$Alive_Dead) ~ 1)
ggsurvplot(fit.surv, data = df, xlab = "Months", ylab = "Estimated Probability of Survival", title = 'Kaplan Meir Survival Curve', ggtheme = theme_minimal()) 

```


```{r}
fit.sex <- survfit(Surv(df$Survival_fromMDM , df$Alive_Dead) ~ df$Gender)


ggsurvplot(fit.sex, data = df, conf.int = TRUE, pval = FALSE, risk.table = FALSE, xlab = "Months", ylab = "Estimated Probability of Survival", title = 'Survival Curve by Gender', legend.labs = c("Female", "Male"), legend = "right", break.time.by = 10, legend.title = "Year", censor.shape = "|", censor.size = 3, palette = c("firebrick", "goldenrod1"), ggtheme = theme_minimal())

logrank.test <- survdiff (Surv(df$Survival_fromMDM , df$Alive_Dead) ~ df$Gender)
logrank.test
```


```{r}
fit.year <- survfit(Surv(df$Survival_fromMDM , df$Alive_Dead) ~ df$Year)
ggsurvplot(fit.year, data = df, conf.int = TRUE, pval = FALSE, risk.table = FALSE, xlab = "Months", ylab = "Estimated Probability of Survival", title = 'Survival Curve by Year', legend.labs = c("pandemic", "pre-pandemic"), legend = "right", break.time.by = 10, legend.title = "Year", censor.shape = "|", censor.size = 3, palette = c("firebrick", "goldenrod1"), ggtheme = theme_minimal())


logrank.test <- survdiff (Surv(df$Survival_fromMDM , df$Alive_Dead) ~ df$Year)
logrank.test

```


```{r}
fit.hcc <- survfit(Surv(df$Survival_fromMDM , df$Alive_Dead) ~ df$HCC_BCLC_Stage)

ggsurvplot(fit.hcc, data = df, conf.int = TRUE, pval = FALSE, risk.table = FALSE, legend.labs=c("0", "A", "B", "C", "D" ),legend = "right", break.time.by = 10, legend.title = "HCC_BCLC_Stage", censor.shape = "|", censor.size = 3, palette=c( "red", "goldenrod1", "skyblue","blue","firebrick" ), xlab = "Months", ylab = "Estimated Probability of Survival",  ggtheme = theme_minimal())


logrank.test <- survdiff (Surv(df$Survival_fromMDM , df$Alive_Dead) ~ df$HCC_BCLC_Stage)
logrank.test
```


```{r}
# Fit survival curves by Mode_Presentation
fit.mode <- survfit(Surv(df$Survival_fromMDM, df$Alive_Dead) ~ df$Mode_Presentation)

# Create ggsurvplot with similar settings
ggsurvplot(
  fit.mode,
  data = df,
  conf.int = TRUE,
  pval = FALSE,
  risk.table = FALSE,
  legend.labs = levels(df$Mode_Presentation),
  legend = "right",
  legend.title = "Mode Presentation",
  palette = c("skyblue", "goldenrod1", "firebrick"),
  xlab = "Months",
  ylab = "Estimated Probability of Survival",
  ggtheme = theme_minimal()
)


logrank.test <- survdiff (Surv(df$Survival_fromMDM , df$Alive_Dead) ~ df$Mode_Presentation)
logrank.test
```
```{r}
km_cancer <- survfit(Surv(df$Survival_fromMDM, df$Alive_Dead) ~ df$Cancer, data = df)
ggsurvplot(km_cancer, data = df, conf.int = TRUE, pval = FALSE, risk.table = FALSE, xlab = "Months", ylab = "Estimated Probability of Survival", title = 'Survival Curve by Year', legend = "right", break.time.by = 10, legend.title = "Cancer Status", censor.shape = "|", censor.size = 3, palette = c("firebrick", "goldenrod1"), ggtheme = theme_minimal())



logrank.test <- survdiff (Surv(df$Survival_fromMDM , df$Alive_Dead) ~ df$Cancer)
logrank.test
```
```{r}
summary(df)
```


```{r}
attach(df)
# Fit the Cox proportional hazards model
cox_model <- coxph(Surv(Survival_fromMDM, Alive_Dead) ~ Month + Year + Age + Size + PS + Cancer + Cirrhosis + Mode_Presentation + Gender + HCC_BCLC_Stage  + HCC_TNM_Stage + Treatment_grps + Prev_known_cirrhosis + Surveillance_programme, data = df)
summary(cox_model)

# Get the survival probabilities from the fitted model
survival_curve <- survfit(cox_model)

```
```{r}
cox_zph <- cox.zph(cox_model)

# Plot Schoenfeld residuals against time for each covariate
plot(cox_zph)
```

```{r}
df <- df[df$Survival_fromMDM > 0, ]

surv_obj = Surv(df$Survival_fromMDM, df$Alive_Dead)
X <- model.matrix(~ . - Survival_fromMDM - Alive_Dead, data = df)



show_fit=glmnet(X,surv_obj,standardize=TRUE,lambda=seq(0,1,.01),family = "cox")
plot(show_fit,label=TRUE)
plot(show_fit,xvar = "lambda",label=TRUE)

surv_model = cv.glmnet(X,surv_obj, family = "cox", type.measure = "deviance",
alpha=1, nfolds = 5)
plot(surv_model)
print(surv_model)
```
```{r}
df_fit <- df[, !(names(df) %in% c("Survival_FDM", "Alive_Dead"))]

lambda_opt = surv_model$lambda.min
fit = glmnet(df_fit, surv_obj, family = "cox", lambda = lambda_opt)
```


```{r}

# Define survival object
surv_obj <- Surv(df$Survival_fromMDM, df$Alive_Dead)

# Remove 'Survival_fromMDM' and 'Alive_Dead' columns from the data frame
df_fit <- df[, !(names(df) %in% c("Survival_fromMDM", "Alive_Dead"))]

# Create model matrix
X <- model.matrix(~ . - 1, data = df_fit)

# Fit Cox model using glmnet
show_fit <- glmnet(X, surv_obj, standardize = TRUE, lambda = seq(0, 1, 0.01), family = "cox")

# Plot the coefficients
plot(show_fit, label = TRUE)

# Plot lambda path
plot(show_fit, xvar = "lambda", label = TRUE)

# Cross-validate glmnet model
surv_model <- cv.glmnet(X, surv_obj, family = "cox", type.measure = "C", alpha = 1, nfolds = 5)

# Plot cross-validation results
plot(surv_model)

# Print cross-validation results
print(surv_model)

# Find optimal lambda
lambda_opt <- surv_model$lambda.min

# Fit final model using optimal lambda
fit <- glmnet(X, surv_obj, family = "cox", lambda = lambda_opt)

# Make predictions
predictions <- predict(fit, newx = X, type = "response")

# Calculate concordance index
c_index <- Cindex(predictions, surv_obj)
print(c_index)
```

